---
title: "Milwaukee Brewers Questionnaire 2022"
author: "Brandon Flores"
output: 
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_folding: "show"
---
\
\
\

### **Problem 1**
Assume that in every inning a starting pitcher pitches he’s projected to give up 0 runs 60% of the time and 1 run 40% of the time. Assume similarly that a relief pitcher is projected to give up 0 runs 70% of the time and 1 run 30% of the time. Assume the starting pitcher will throw 165 innings and the relief pitcher will throw 50 innings in a season. What are the odds that the starting pitcher gives up less runs on a per-inning basis than the relief pitcher?

```{r echo=TRUE}
#starting pitcher simulation
sp <- rbinom(n = 1000000, size = 165, prob = 0.4) 
sp <- sp / 165

#relief pitcher simulation

rp <- rbinom(n = 1000000, size = 50, prob = 0.3)
rp <- rp / 50

#result

result <- sp < rp
num <- (sum(result, na.rm = TRUE))/1000000

odds <- num/(1-num)

```
The probability of that the starting pitcher outperforms the relief pitcher on a per-inning basis is `r num*100`%. The odds are therefore **`r odds`/1**. 
\
\
\

### **Problem 2**
A pitcher throws 60% of his fastballs in the strike zone. Given that a pitch is in the strike zone, the home plate umpire is known to call a strike 75% of the time, and given a pitch out of the strike zone he is known to call a ball 65% of the time. A fastball from this pitcher in a game with the umpire behind the plate is called a ball. What is the probability that that pitch was in the strike zone?
\
\
\

### **Problem 3**
A linear regression model with a normal error term is fit to predict ERA as a function of K% (strikeouts per plate appearance) and BB% (walks per plate appearance) for pitchers. The estimated coefficients on the linear model are as follows:
\
Intercept: 5.81 
\
Slope on K%: -11.7 
\
Slope on BB%: 11.4 
\
Residual standard error/standard deviation of error on estimates: 1.31 

a) What is the probability that a pitcher with a K% of 0.20 and a BB% of 0.10 outperforms a pitcher with a K% of 0.25 and a BB% of 0.20 under this model?
```{r}
#Predicted ERA values for each pitcher based on model
Pitcher1_ERA <- 5.81 - (11.7 * 0.2) + (11.4 * 0.1) # = 4.61
Pitcher2_ERA <- 5.81 - (11.7 * 0.25) + (11.4 * 0.2) # = 5.165

#What is the probability that pitcher 1's ERA is lower than pitcher 2's ERA under this model?

prob <- pnorm(5.164, mean = 4.61, sd = 1.31)
```

Under this model, the pitcher with K% = 0.2 and BB% = 0.1 will outperform the pitcher with K% = 0.25 and BB% = 0.20 approximately **`r prob*100`%** of the time. 

b) If Hit by Pitches per PA were to be added to this model, would you expect the coefficient on BB% to be greater or lower in magnitude? Why? 
\
\
I would expect the coefficient for BB% to go down if Hit By Pitches per PA were added
since BB% is most likely directly correlated to ERA as well as the fact that HBP/PA
is a subset of BB%. This would split the effect that BB% has on ERA into BB% and HBP%.
We can probably safely assume that HBP/PA is also directly correlated to ERA, and would
therefore be a positive coefficient. Therefore, if the coefficient for BB% were to go
up with the addition of HBP/PA, then...



c) On the same data set, a logistic regression model is fit with the dependent variable being 1 if ERA is less than 4.00 and 0 if ERA is greater than 4.00. The estimated coefficients on the logistic regression model are as follows: Intercept: -2.18 K%: 15.9 BB%: -16.1 Is the marginal effect of K% on the probability of having an ERA < 4.00 larger for someone with a K% of 0.20 and BB% of 0.10 or a pitcher with a K% of 0.25 and BB% of 0.20?

```{r}
fun <- deriv(expression((exp(-2.18+(15.9*x1)-16.1*x2))/(1+exp(-2.18+(15.9*x1)-16.1*x2))), c("x1", "x2"), function.arg = TRUE)
fun(0.2, 0.1)
fun(0.25, 0.2)
```

The marginal effect of K% on the probability of a pitcher's ERA being less than 4.00
is greater for the pitcher with K% = 0.2 and BB% = 0.1. This implies that the first
pitcher's ERA would benefit more from increasing his K% than the second pitcher based
on this model.
\

d) If forced to choose between these linear or logistic models based on K% and BB%, which would
you prefer to use to determine the quality of a pitcher from only their K% and BB%? Why?
\
\
\

### **Problem 4**
Entering the day of September 26, 2022 the New York Mets had a record of 97-57 and the Atlanta
Braves were 95-58. In their remaining games, the Mets and Braves were set to play each other in a
3-game series, plus each had 2 other series against other opponents. To that point in the season
the Mets’ record against the Braves was 9-7, and if the two teams tied in overall record at the end
of the season, whichever team won the season series against each other would win the division.
Assume that the Mets’ odds of winning each of their remaining games was 50%. Assume that the
Braves’ odds of winning each game against the Mets was 50% and p against all non-Mets
opponents. For what value of p would give the Braves a 50% chance of winning the division?

```{r message=FALSE}
library(tidyverse)

braves_mets_sim <- function(p){
  braves_mets <- rbinom(1000000, 3, 0.5)
  mets_braves <- 3 - braves_mets

  braves_other <- rbinom(1000000, 6, p)
  mets_other <- rbinom(1000000, 5, 0.5)

  braves_total <- braves_mets + braves_other + 95
  mets_total <- mets_other + mets_braves + 97

  braves_win_div <- (braves_total > mets_total) | ((braves_total == mets_total) & braves_mets == 3)

  return(sum(braves_win_div, na.rm = TRUE) / 1000000)
}

range <- seq(0.5, 0.85, by = 0.05)
range <- seq(0.8, 0.84, by = 0.001)

values <- map_dbl(.x = range, .f = ~braves_mets_sim(.x))



plot(range, values)
```

\
\
\

### **Problem 5**
Assume that a newly drafted player has WAR/S projections (WAR/S defined as Wins Above
Replacement per Season, where the scale of a Season is defined as 600 Plate Appearances) in each
of his first 12 professional seasons distributed according to a normal distribution with mean and
standard deviation parameters provided below.
Assume that in each season his expected MLB-level playing time is a calculated as a function of his
projection, weighted by the probability of the given projection according to the stated probability
distribution. For a given WAR/S projection, p, the playing time function is:

$$
PlayingTime (PA) = \left\{\begin{aligned}
&0 &&    p < 0\\
&600*(p/2) && 0 < p \le 2\\
&600 && p > 2
\end{aligned}
\right.$$

What is this player’s total expected MLB-level playing time over the 7-year period where the
projected cumulative playing time is at a maximum?

```{r include=FALSE}
library(knitr)
library(kableExtra)
```

```{r echo=FALSE}
season <- c(1:12)
WAR_per_S <- c(-2.6, -1.5, -0.5, 0.2, 0.8, 1.4, 1.5, 1.4, 1.2, 0.9, 0.5, -0.1)
sd <- c(0.2, 0.7, 1.0, 1.3, 1.5, 1.7, 1.9, 2.1, 2.3, 2.4, 2.6, 2.8)
df <- data.frame(season, WAR_per_S, sd)
kable(df, col.names = c("Season", "Mean WAR/S Projection","StandardDeviation"), align = c("c", "c", "c")) %>%
  kable_styling(bootstrap_options = "bordered")

```

```{r}
library(tidyverse)


#Calculating P(WAR-0.05 < X < WAR+0.05)
df <- df %>% mutate(prob = pnorm(WAR_per_S+0.5, mean = WAR_per_S, sd = sd)-pnorm(WAR_per_S-0.05, mean = WAR_per_S, sd = sd)) 

playing_time <- function(p){
  if(p < 0){
    return(0)
  }else if(p>0 & p<=2){
    return(600*(p/2))
  }else{
    return(600)
  }
}

df <- df %>% 
  rowwise() %>%
  mutate(PA_unweighted = playing_time(WAR_per_S))

df <- df %>%
  mutate(PA_weighted = PA_unweighted * prob)
View(df)

#Looking at the data frame we can see that the maximum playing time occurs during 
# seasons 5 to 11

answer <- sum(df$PA_weighted) # = 264.0842 PA
```
The expected cumulative playing time is estimated to be **`r answer` Plate Attempts**. 
\
\
\

### **Problem 6**
Given the following information, describe how you would construct a model to predict the likelihood of a flyball out on a flyball to the outfield:
Identities of all players on field (fielders, pitcher, batter, runners), game state (number of outs, runners on base, inning, score), temperature, crowd size, batted ball launch characteristics (exit velocity, launch angle, spray angle, hang time), outfielder fielding percentage, stadium characteristics (fence distance and height), day/night game, pitcher/batter historical matchup results, initial fielder positioning data.
Note: There is no need to find data and generate an actual model. Please just describe how you would proceed assuming you had the data listed. Please limit your response to 300 words.


\
\
\

### **Problem 7**
The Brewers are considering signing one of two different infielders to 4-year guaranteed contracts with WAR/S projections and salaries in the following table:
```{r echo=FALSE}
Season <- c(2023,2023, 2024, 2024, 2025, 2025, 2026, 2026)
Player <- c("Player A", "Player B", "Player A", "Player B", "Player A", "Player B", "Player A", "Player B")
WAR_s <- c(1.1, 3.1, 1.7, 2.4, 1.5, 1.6, 1.2, 0.7)
Salary <- c(1.0, 10.0, 3.0, 10.0, 7.0, 10.0, 10.0, 10.0)
df <- data.frame(Season, Player, WAR_s, Salary)
dfA <- df %>% filter(Player == "Player A") %>%
  select(-Player)
knitr::kable(dfA, caption = "Player A", col.names = c("Season", "WAR/S","Salary ($M)"), align = c("c", "c", "c")) %>%
  kable_styling(bootstrap_options = "bordered")
dfB <- df %>% filter(Player == "Player B") %>%
  select(-Player)
knitr::kable(dfB, caption = "Player B", col.names = c("Season", "WAR/S","Salary ($M)"), align = c("c", "c", "c")) %>%
  kable_styling(bootstrap_options = "bordered")
```

Assuming you had to pick one of these two deals, which player would you recommend the Brewers sign. Note: there are no forfeited draft picks or any other compensation required for the contracts. Please explain your process and any assumptions you make. Please limit your answer to 300 words

\
\
\

### **Problem 8**
In 2022, two minor leagues used an automated ball-strike (ABS) challenge system, in which the umpire calls balls and strikes, but the pitcher, catcher, and batter may challenge the call. Each team is allocated three challenges per game with successful challenges retained. If this system were to be implemented in the majors in 2024, how should teams adjust their team building and gameplay strategies to prepare? Please limit your answer to 300 words.